|

### 2022-12-20 paper_comparison/paper_comparison/paper_comparison.Rmd
update before meeting

(discuss choice of tests, mention pkg `lmtest`, `skedastic`)

@@

To compare the power of visual test and conventional tests, we picked RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`. These three tests can be used to detect non-linearity, heteroskedasticity and non-normality respectively. Their estimated power is shown in Figure \ref{fig:polypower}. For the RESET test, we included different power of fitted values as proxies. According to [@ramsey_tests_1969], there is no general rule for the power of the fitted values needed by the RESET test, but it found power up to four is usually sufficient. Thus, we followed this guideline to conduct the RESET test. For the Breusch-Pagan test, both $\boldsymbol{x}$ and $\boldsymbol{x}^2$ were included in the auxiliary regression equation to [justify this choice?].

@@

For the non-linearity model, the power of all four tests increases as the effect gets larger. The power curve of RESET test climbs aggressively from 13% to around 70% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect and remain lower than 25% throughout the period, showing that F-test is way more sensitive to the type of model defects that being considered. Meanwhile, no noticeable visual features can be spotted from the example residuals plots.

@@

Overall, the power analysis suggests that conventional tests do not match how humans perceive residual plots and thus determine model violations in these two scenarios. Non-severe model violations with human acceptable residual departures could often alarm by the conventional tests. This could limit the popularity of conventional tests in residual diagnostics among analysts.

@@

```{r p-value-comparison, fig.height=4, fig.cap="Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model. The scatter plot is drawn on a square root scale. Every dot on the plot corresponds to a lineup. Most of the dots are fallen on the left of the y = x line indicating visual test p-value is almost always higher than conventional test p-value. The blue curve is a smoothing of the dots showing a positive trend. "}
vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  left_join(vi_survey %>% 
           count(unique_lineup_id) %>%
           rename(num_subject = n)) %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  ggplot() +
  geom_abline() +
  geom_point(aes(conventional_p_value, p_value), alpha = 0.4) +
  geom_smooth(aes(conventional_p_value, p_value), se = FALSE, span = 0.6) +
  scale_x_sqrt() +
  scale_y_sqrt() +
  facet_wrap(~type) +
  ylab("Visual test p-value on square root scale") +
  xlab("Conventional test p-value on square root scale") +
  theme_light()
```

>>>>

@@

To compare the power of visual test and conventional test, we picked RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`. Among them, RESET test is the only exact and appropriate test in this scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests. Their estimated power is shown in Figure \ref{fig:polypower}. For the RESET test, we included different power of fitted values as proxies. According to @ramsey_tests_1969, there is no general rule for the power of the fitted values needed by the RESET test, but it found power up to four is usually sufficient. Thus, we followed this guideline to conduct the RESET test. For the Breusch-Pagan test, both $\boldsymbol{x}$ and $\boldsymbol{x}^2$ were included in the auxiliary regression equation to [justify this choice?].

@@

For the non-linearity model, the power of all four tests increases as the effect gets larger. The power curve of RESET test climbs aggressively from 13% to around 70% as $log_e(E)$ increases from 0 to 2, while power of other tests respond inactively to the change of effect and remain lower than 25% throughout the period, showing that REST test is way more sensitive to the type of model defects that being considered. Meanwhile, no noticeable visual features can be spotted from the example residuals plots.

@@

Overall, the power analysis suggests that conventional tests do not match how humans perceive residual plots in these two scenarios. Humans have much higher tolerance of the residual departures than the conventional test. Since fail to reject $H_0$ in a visual test means that there is no obvious visual discoveries found in the residual plot, analysts and the general public as the consumers of the output may not be convinced of the existence of significant residual departures in spite of the rejection of $H_0$ given by the conventional test. Even if the rejection is accepted, the model violation may be considered as impactless due to the fact that they are not clearly visible. Besides, the sensitivity of the conventional test could also distract and discourage analysts from finding simple but good linear approximation to the data. The rejection of $H_0$ because of human acceptable and negligible residual departures is not practically meaningful and useful. (it does not fit the purpose of )

This could limit the popularity of conventional tests in residual diagnostics among analysts.

@@

```{r p-value-comparison-old, eval = FALSE, fig.height=4, fig.cap="Visual test p-value compared to conventional test p-value for lineups produced by heteroskedasticity and non-linearity model. The scatter plot is drawn on a square root scale. Every dot on the plot corresponds to a lineup. Most of the dots are fallen on the left of the y = x line indicating visual test p-value is almost always higher than conventional test p-value. The blue curve is a smoothing of the dots showing a positive trend. "}
vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  left_join(vi_survey %>% 
           count(unique_lineup_id) %>%
           rename(num_subject = n)) %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  ggplot() +
  geom_abline() +
  geom_point(aes(conventional_p_value, p_value), alpha = 0.4) +
  geom_smooth(aes(conventional_p_value, p_value), se = FALSE, span = 0.6) +
  scale_x_sqrt() +
  scale_y_sqrt() +
  facet_wrap(~type) +
  ylab("Visual test p-value on square root scale") +
  xlab("Conventional test p-value on square root scale") +
  theme_light()
```

```{r p-value-comparison, fig.height=4, fig.cap="Rejection rate of visual test compared to conventional test on lineups produced by heteroskedasticity and non-linearity model. Based on the rejection decision made by the visual test and the conventional test, the mosaic plot shows the proportion of lineups in each category."}

library(ggmosaic)
vi_survey %>%
  filter(!null_lineup) %>%
  filter(!attention_check) %>%
  mutate(type = ifelse(type == "polynomial", "non-linearity", type)) %>%
  left_join(vi_survey %>% 
           count(unique_lineup_id) %>%
           rename(num_subject = n)) %>%
  group_by(unique_lineup_id) %>%
  summarise(across(everything(), first)) %>%
  mutate(conv_reject = ifelse(conventional_p_value < 0.05, "Reject", "Not reject"),
         reject = ifelse(p_value < 0.05, "Reject", "Not reject")) %>%
  ggplot() +
  geom_mosaic(aes(x = product(conv_reject, reject), fill = conv_reject)) +
  scale_fill_manual(values = rev(rcartocolor::carto_pal(3, "Vivid"))[2:3]) +
  xlab("Visual test") +
  ylab("Conventional test") +
  labs(fill = "Conventional test")
```
